Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-11-26 manuscript/manuscript.qmd
Add comments by NSG

targets::tar_load(
	c(
		participants,
		stimuli,
		vocabulary,
		attrition_trials,
		attrition_participants,
		bvq_data,
		gaze,
		data_bcn,
		data_oxf
	)
)

targets::tar_load(
	c(
		model_fits_bcn,
		model_loos_bcn,
		model_fits_oxf,
		model_loos_oxf
	)
)

# prepare data
attrition <- lst(
	participants,
	attrition_participants,
	attrition_trials
) |>
	reduce(inner_join, by = join_by(session_id)) |>
	unnest_wider(matches("all"), names_sep = "_") |>
	rename_with(\(x) gsub("is_valid_gaze_all_|is_valid_vocab_all", "", x))

@@

# set ggplot theme and colour palette
my_theme <- theme_minimal() +
	theme(
		panel.grid = element_blank(),
		axis.line = element_line(colour = "black"),
		text = element_text(size = 12, colour = "black"),
		axis.text = element_text(colour = "black")
	)

@@

The implicit naming paradigm provides an ideal experimental paradigm study the developing bilingual lexicon. By covertly manipulating the cross-linguistic relationship between the prime and target labels, parallel activation can be tested while participants are presented with auditory stimuli (target labels) exclusively in one of their languages [see @von2014bilinguals for a similar approach in bilingual adults]. Capitalizing on the language non-selective account of lexical access, we exploited the implicit naming to investigate the mechanisms behind the emergence of phonological priming effects in the bilingual developing lexicon. We tested a cohort of monolingual and bilingual infants learning Catalan and Spanish between 20 and 32 months of age. This allowed us to compare the performance of participants with differing vocabulary sizes in the word recognition task. In order to circumvent the problem of limited vocabulary knowledge in the non-dominant language, we tested participants only in their dominant language [@costa2014does].

Following @mani2010infant, each trial in the task started with the silent presentation of a prime picture. Both monolingual and bilingual infants were expected to implicitly name the prime picture. According to the language non-selective hypothesis of lexical access, bilinguals should generate two labels for the prime picture, one in each language. To test this prediction, we manipulated the phonological similarity between the prime and the target words in both languages (see @fig-hypotheses). In *Related/Non-cognate* trials, prime and target labels shared phonological onset in only the language of test. For instance an infant tested in Catalan would be presented with a chair as prime picture ($\textbf{c}\text{adira}_{\text{CAT}}/\text{silla}_{\text{SPA}}$) and with $\textbf{c}\text{ullera}_{\text{CAT}}$ [spoon] as target label. In line with @mani2011phonological, we anticipated that the phonological overlap between prime and target should modulate target word recognition in both monolinguals and bilinguals. This should be reflected in a delayed target looking preference, compared to *Unrelated* trials, in which prime and target did not share phonological onset. In *Related/Cognate* trials, the prime shared phonological onset with the target in both languages. For instance, the same infants tested in Catalan would be presented with a car as prime picture ($\textbf{c}\text{otxe}_{\text{CAT}}/\textbf{c}\text{oche}_{\text{SPA}}$) and with $\textbf{c}\text{ullera}_{\text{CAT}}$ [spoon] as target label. In bilinguals, parallel activation of the prime in both languages should increase the cohort of the target word, leading to stronger interference effects in this condition, compared to *Related/Non-cognate* and *Unrelated* trials.

@@

In line with previous studies in monolinguals, we further predicted that the strength of the lexical interference effects in the *Related/Non-cognate* and *Related/Cognate* conditions would be associated to participants' vocabulary size. Target word recognition should be delayed by the activation of a larger cohort of phonologically related words [@mani2011phonological; @chow2017spoken]. In the case of bilinguals, two measures of vocabulary size are relevant: *total vocabulary size* and *dominant-language vocabulary size*. Total vocabulary size refers to the number of words that a child has acquired, regardless of the language they belong to. Dominant-language vocabulary size (L1 vocabulary size, henceforth) refers to the number of acquired words in the dominant language of the child (in which they were tested), the language in which they know more words. If the size of interference effects are more tightly associated to bilinguals’ total vocabulary size, it would support the assumption that lexical representations from both languages enter a unique cohort in bilinguals.

@@

@@

```{r}
#| label: participants-numbers-oxf
participants_oxf <- filter(participants, location == "Oxford")

@@

	cols_merge(c(age_min, age_max), pattern = "{1}-{2}") |>
	cols_merge(c(age_mean, age_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_catalan_mean, doe_catalan_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_english_mean, doe_english_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_spanish_mean, doe_spanish_sd),
			   pattern = "{1} ({2})") |>
	tab_spanner("Sample size", starts_with("n")) |>
	tab_spanner("Age (months)", matches("age")) |>
	tab_spanner("Catalan", matches("catalan")) |>
	tab_spanner("English", matches("english")) |>
	tab_spanner("Spanish", matches("spanish")) |>
	tab_spanner("Degree of Exposure (%)", matches("doe")) |>
	cols_label(n = "Sessions",
			   n_participants = "Participants",
			   age_mean = "M (SD)",
			   age_min = "Range",
			   doe_catalan_mean = "M (SD)",
			   doe_english_mean = "M (SD)",
			   doe_spanish_mean = "M (SD)") |>
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	tab_style(cell_text(style = "italic"),
			  cells_column_spanners())
```

@@

We collected vocabulary data using parental responses to the Oxford Communicative Development Inventory (OCDI) [@hamilton2000infant]. Families were sent the questionnaire immediately after each experimental session, and were given two weeks to fill it. In the case that a complete response to the OCDI was not provided within the two-week limit, the participants' testing session was excluded from the analyses (*n* = `r vocabulary_oxf_exc`). @fig-vocabulary-oxf shows the distribution of participants' vocabulary sizes across ages.

```{r fig-vocabulary}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles. Vocabulary size scores from the recurrent participants are shown linked. For visualisation purposes, testing sessions were aggregated at three age points: 21, 25, and 30 months. Points and intervals indicate mean and standard deviations, respectively."
#| fig-width: 6
#| fig-height: 4
vocab_plot_data <- vocabulary |>
	inner_join(filter(attrition_participants, is_valid_participant),
			   by = join_by(session_id)) |>
	inner_join(select(participants, location, lp, 
					  child_id, session_id, age_group),
			   by = join_by(child_id, session_id)) |>
	pivot_longer(c(l1_prop, total_prop),
				 names_to = "type",
				 values_to = "prop") |> 
	mutate(location = factor(location, levels = c("Oxford", "Barcelona")),
		   lp = gsub(" \\(English\\)", "", lp),
		   side = case_when(location=="Oxford" ~ "both",
		   				 lp=="Monolingual" ~ "left",
		   				 lp=="Bilingual" ~ "right"),
		   type = if_else(type=="l1_prop", 
		   			   "L1 vocabulary", "Total vocabulary"),
		   study = if_else(location=="Oxford", 
		   				"Study 1 (English)", "Study 2 (Catalan/Spanish)"))

vocab_plot_means <- vocab_plot_data |> 
	summarise(prop_sd = sd(prop, na.rm = TRUE),
			  prop = mean(prop, na.rm = TRUE),
			  n = n(),
			  .by = c(age_group, type, side, lp, type, location)) |> 
	mutate(.lower = prop - (prop_sd/sqrt(n)),
		   .upper = prop + (prop_sd/sqrt(n)))

vocab_plot_data |> 
	ggplot(aes(age_group, prop, 
			   colour = lp,
			   fill = lp,
			   side = side)) +
	facet_grid(type~study) +
	geom_line(aes(group = child_id),
			  colour = "grey",
			  alpha = 1 / 4, linewidth = 3 / 4) +
	geom_dots(aes(side =),
			  dotsize = 1.5,
			  scale = 0.25,
			  binwidth = 0.025,
			  alpha = 2 / 4) +
	geom_errorbar(data = vocab_plot_means,
				  aes(ymin = .lower,
				  	ymax = .upper),
				  colour = "black",
				  show.legend = FALSE,
				  width = 0.15,
				  linewidth = 1/2,
				  position = position_dodge(width = 1)) +
	geom_point(data = vocab_plot_means, 
			   colour = "black",
			   show.legend = FALSE,
			   size = 1.25,
			   position = position_dodge(width = 1)) +
	# geom_label(data = vocab_plot_means,
	# 		   aes(label = sprintf("%.1f%%\n[%.1f, %.1f]",
	# 		   					prop*100, .lower*100, .upper*100),
	# 		   	y = 0),
	# 		   size = 2.5,
	# 		   fill = "white",
	# 		   vjust = 0,
	# 		   show.legend = FALSE,
	# 		   label.r = unit(0, "lines"),
	# 		   label.size = 0,
	# 		   label.padding = unit(0.1, "lines"),
# 		   position = position_dodge(1),
# 		   alpha = 1/4) +
# geom_point(alpha = 1) +
labs(x = "Age (months)",
	 y = "Vocabulary size",
	 colour = "Group",
	 fill = "Group") +
	scale_y_continuous(labels = scales::percent,
					   limits = c(0, 1)) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  axis.title.x = element_blank(),
		  legend.title = element_blank())
```

@@

Participants were presented with 32 trials in random order, which belonged to three conditions: *Unrelated* (16 trials), *Related/Non-cognate* (8 trials), and *Related/Cognate* (8 trials). In *Unrelated* trials, target was phonologically unrelated with the English and Spanish labels of the prime picture (e.g., $\text{door}_{\text{ENG}}/\text{puerta}_{\text{SPA}}-\text{sock}_{\text{ENG}}$). In *Related/Non-cognate* trials, the target shared phonological onset with the English prime label, but not with the Spanish prime label (e.g., $\textbf{t}\text{ree}_{\text{ENG}}/\text{árbol}_{\text{SPA}}-\textbf{t}\text{ruck}_{\text{ENG}}$). In *Related/Cognate* trials, the target shared phonological overlap with both English and Spanish prime labels ($\textbf{b}\text{oot}_{\text{ENG}}/\textbf{b}\text{ota}_{\text{SPA}}-\textbf{b}\text{ird}_{\text{ENG}}$). Target-distractor pairs were phonologically unrelated, and held constant for all participants. Especial attention was paid to avoiding semantic or taxonomic relationships between prime and target words, and between prime and distractor words. Each target distractor pair was shown twice for each participant, switching roles. Distractors were always phonologically unrelated to the prime and target labels in the same trial.

We created fours lists of trials, across which the same target-distractor pair appeared with a different prime, counterbalancing the condition to which it belonged. For instance, in list one the *ball*-*trousers* pair was preceded by *bike* (*Related/Cognate*), by *butterfly* (*Related/Non-cognate*) in list two, and by *star* and *nose* (*Unrelated*) in lists three and four. @tbl-stimuli shows a detailed summary of the stimuli properties, broken down by trial type and testing language. Lexical frequencies were extracted from the English corpora from the CHILDES database [@macwhinney2000childes; @sanchez2019childes] as counts per million words, and transformed into Zipf scores for easier cross-language comparison [@zipf1945meaning; @van2014subtlex].

```{r tbl-stimuli}
#| label: tbl-stimuli
#| tbl-cap: "Summary of stimuli properties by trial type."
stimuli |>
	unnest_wider(c(freq, nphon)) |>
	summarise(across(c(matches("n_phon|freq_|lv_|nphon"), duration),
					 list(	mean = \(x) mean(x, na.rm = TRUE),
					 	  sd = \(x) sd(x, na.rm = TRUE))),
			  .by = c(trial_type, test_language)) |>
	select(-matches("distractor")) |>
	mutate(test_language = if_else(test_language=="English",
								   paste0("Study 1: ", test_language),
								   paste0("Study 2: ", test_language))) |>
	arrange(test_language) |> 
	gt(rowname_col = "trial_type",
	   groupname_col = "test_language") |>
	# summary_rows(groups = c("Unrelated", "Non-cognate", "Cognate"),
	# 			 columns = is.numeric,
	# 			 fns = list(Total ~ mean(.)),
	# 			 fmt = ~ fmt_number(., decimals = 2)) |>
	grand_summary_rows(	columns = is.numeric,
						fns = list(Total ~ mean(.)),
						fmt = ~ fmt_number(., decimals = 2)) |>
	tab_spanner("Phonemes", matches("nphon")) |>
	tab_spanner("Frequency (Zipf)", matches("freq")) |>
	tab_spanner("Phon. similarity [0-1]", matches("lv")) |>
	fmt_number(is.numeric, decimals = 1) |>
	cols_merge(c(freq_prime_mean, freq_prime_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(freq_target_mean, freq_target_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(nphon_prime_mean, nphon_prime_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(nphon_target_mean, nphon_target_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(lv_pp_mean, lv_pp_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(lv_pt_mean, lv_pt_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(duration_mean, duration_sd),
			   pattern = "{1} ({2})") |>
	cols_label(trial_type = "",
			   freq_prime_mean = "Prime",
			   freq_target_mean = "Target",
			   lv_pp_mean = "Prime-Prime",
			   lv_pt_mean = "Prime-Target",
			   duration_mean = "Duration (s)",
			   nphon_prime_mean = "Prime",
			   nphon_target_mean = "Target") |>
	tab_style(cell_text(style = "italic"),
			  list(cells_column_labels(),
			  	 cells_column_spanners(),
			  	 cells_grand_summary()))
```

The auditory stimuli were natural exemplars of the selected target words, spoken by a Southern British English female speaker who was instructed to pronounce each word in a toddler-directed manner. We used the Audacity and Praat [@boersma2001speak] software packages to trim, denoised, and normalise their amplitude. The visual stimuli were realistic photographic representations of a typical exemplars of the prime, target, and distractor words. Image backgrounds were removed from the original pictures using the GNU Image Manipulation Program (GIMP), resized to a rectangle of a maximum of 400 pixels hight or wide, and finally placed in the centre of a 50% grey rectangle square of 500 $\times$ 500 pixels. The final stimuli had a resolution of 72 dpi. When presented in the eye-tracker screen, the areas of interest (AOI) occupied an area of 13.23 $\times$ 13.23 cm (11.613$^{\circ}$ visual angle from participants' perspective).

### Procedure and apparatus

Testing took place in a sound-proof room. Participants sat on their caregivers' lap in a dimly lit testing booth while the experimenter conducted the experiment from outside. Caregivers were instructed to keep their eyes shut (to avoid recording their gaze, instead of the participant's), to be still, and to avoid interacting with the participant verbally or non-verbally. Participants sat at approximately 65 cm from the eye-tracker and a XX-in screen of $1929\times1080$ screen resolution. The study was run using a custom Matlab script, PresentMate, based on the PsychToolbox-3 extension [@kleiner2007s; @brainard1997psychophysics; @pelli1997videotoolbox]. Visual fixations were recorded using a Tobii TX 300 eye-tracker and a 23-in screen of 1920 $\times$ 1080 resolution. The Tobii Analytics SDK 3.0 was used to interact with the eye-tracking while the experiment was running. Sampling rate was set at 120 Hz. A 9-point calibration was performed before every experimental session, in which the picture of a colourful beach ball was presented. We set a 55% grey background for the screen during calibration and stimuli presentation. Auditory stimuli were presented through two loudspeakers located behind the screen, one to each side. The experimenter monitored the experimental from outside the room using a centrally located video camera place above the screen. After a successful calibration the experimenter triggered the onset of the first trial. Trials were presented uninterruptedly and without intervention of the experimenter until the 32 trials were presented, or the experimental session had to be stopped because of the participant's behaviour.

@fig-task shows a illustrates of the task design. Each trial started with the presentation of an attention getter for 3,000 ms. Then, the prime picture was presented in silence in the centre of the screen for 1,500 milliseconds. Fifty milliseconds after the offset of the prime image, an auditory label was played from the loudspeakers and, 700 milliseconds after the onset of the auditory label, the target and distractor pictures were presented side-by-side during 1,000 milliseconds until the end of the trial. After this, the attention getter of the next trial was immediately presented. Each experimental session took approximately 10 minutes.

@@

```{r dataset_oxf}
attrition_oxf <- filter(attrition, location == "Oxford")
n_trials <- nrow(attrition_oxf)
n_sessions <- n_distinct(attrition_oxf$session_id)
n_participants <- n_distinct(attrition_oxf$child_id)
n_trials_valid <- nrow(filter(
	attrition_oxf,
	is_valid_participant,
	is_valid_trial
))
n_sessions_valid <- attrition_oxf |>
	distinct(child_id, .keep_all = TRUE) |>
	pull(is_valid_participant) |>
	sum()

@@

model_summary <- get_posterior_summary(model_fits_bcn$fit_3,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval
)

>>>>

targets::tar_load(c(participants,
					stimuli,
					vocabulary,
					attrition_trials,
					attrition_participants,
					bvq_data,
					gaze,
					data_bcn,
					data_oxf))

targets::tar_load(c(model_fits_bcn,
					model_loos_bcn,
					model_fits_oxf,
					model_loos_oxf))

# prepare data
attrition <- lst(participants,
				 attrition_participants,
				 attrition_trials) |>
	reduce(inner_join, by = join_by(session_id)) |>
	unnest_wider(matches("all"), names_sep = "_") |>
	rename_with(\(x) gsub("is_valid_gaze_all_|is_valid_vocab_all",
						  "", x))

@@

# set ggplot theme and colour palette
my_theme <- theme_minimal() +
	theme(panel.grid = element_blank(),
		  axis.line = element_line(colour = "black"),
		  text = element_text(size = 12, colour = "black"),
		  axis.text = element_text(colour = "black"))

@@

The implicit naming paradigm provides an ideal experimental paradigm to study the developing bilingual lexicon. By covertly manipulating the cross-linguistic relationship between the prime and target labels, parallel activation can be tested while participants are presented with auditory stimuli (target labels) exclusively in one of their languages [see @von2014bilinguals for a similar approach in bilingual adults]. Capitalizing on the language non-selective account of lexical access, we exploited the implicit naming to investigate the mechanisms behind the emergence of phonological priming effects in the bilingual developing lexicon. We tested a cohort of monolingual and bilingual infants learning Catalan and Spanish between 20 and 32 months of age. We compared the performance of participants with differing vocabulary sizes in the word recognition task. In order to circumvent the problem of limited vocabulary knowledge in the non-dominant language, we tested participants only in their dominant language [@costa2014does].

Following @mani2010infant, each trial in the task started with the silent presentation of a prime picture. Both monolingual and bilingual infants were expected to implicitly name the prime picture. According to the language non-selective hypothesis of lexical access, bilinguals should generate two labels for the prime picture, one in each language. To test this prediction, we manipulated the phonological similarity between the prime and the target words in both languages (see @fig-hypotheses). In *Related/Non-cognate* trials, prime and target labels shared phonological onset in only the language of test. For instance an infant tested in Catalan would be presented with a chair as prime picture ($\textbf{c}\text{adira}_{\text{CAT}}/\text{silla}_{\text{SPA}}$ [chair]) and with $\textbf{c}\text{ullera}_{\text{CAT}}$ [spoon] as target label. In line with @mani2011phonological, we anticipated that the phonological overlap between prime and target should modulate target word recognition in both monolinguals and bilinguals. This should be reflected in a delayed target looking preference, compared to *Unrelated* trials, in which prime and target did not share phonological onset. In *Related/Cognate* trials, the prime shared phonological onset with the target in both languages. For instance, the same infants tested in Catalan would be presented with a car as prime picture ($\textbf{c}\text{otxe}_{\text{CAT}}/\textbf{c}\text{oche}_{\text{SPA}}$) and with $\textbf{c}\text{ullera}_{\text{CAT}}$ [spoon] as target label. In bilinguals, parallel activation of the prime in both languages should increase the cohort of the target word, leading to stronger interference effects in this condition, compared to *Related/Non-cognate* and *Unrelated* trials.

@@

In line with previous studies in monolinguals, we further predicted that the strength of the lexical interference effects in the *Related/Non-cognate* and *Related/Cognate* conditions would be associated to participants' vocabulary size. Target word recognition should be delayed by the activation of a larger cohort of phonologically related words [@mani2011phonological; @chow2017spoken]. We defined vocabulary size as the amount of words participants were reported to understand in their dominant language by their caregivers. The choice of the dominant language for calculating vocabulary sizes is due to several reasons. First, it allows a more fair comparison between monolinguals (who do not know any language other than their dominant language) and bilinguals (who may know words in a second language). Second, since participants were tested exclusively in their dominant language, their vocabulary size in the dominant language is more likely to be associated to participants' performance in the task. Third, previous work on word recognition in bilinguals suggests that vocabulary size in the dominant language predicts participants' performance better than total vocabulary (in which vocabulary sizes in both languages are summed together) [@marchman2010vocabulary].

@@

In this study, we conducted a conceptual replication of @mani2010infant and @mani2011phonological. We tested a group of English monolinguals aged 20 to 32 months, living in the Oxfordshire area (United Kingdom). As just said, participants were tested exclusively in English. As in the original studies, we manipulated the phonological relationship between the prime and the target label. We expected participants’ target looking to change as a function of the phonological relatedness between the prime and target English labels. This would reveal that participants implicitly named the prime pictures, generating a phonological label that influenced the subsequent recognition of a phonologically related word. In half of the trials the English prime label was phonologically related to its Spanish translation, that it they were cognates; in the other half they were not phonologically related (non-cognates)^[It was initially planned to collected data from English monolinguals and English-Spanish bilinguals in Oxford, therefore the manipulation of the cognate status of the words in English and Spanish. Due to time limitations imposed by the COVD-19 lockdown between 2020 and 2022, collecting data from bilinguals was not possible. We report the available data from English monolinguals as a control for Catalan and Spanish monolinguals and bilinguals from Barcelona in Study 2.]. Given participants’ lack of knowledge of Spanish (or any language other than English), participants’ performance was predicted to be unaffected by the cognate status of the primes.

@@

```{r}
#| label: participants-numbers-oxf
participants_oxf <- filter(participants, location=="Oxford")

@@

	cols_merge(c(age_min, age_max), pattern = "{1}-{2}") |>
	cols_merge(c(age_mean, age_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_catalan_mean, doe_catalan_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_english_mean, doe_english_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_spanish_mean, doe_spanish_sd),
			   pattern = "{1} ({2})") |>
	tab_spanner("Sample size", starts_with("n")) |>
	tab_spanner("Age (months)", matches("age")) |>
	tab_spanner("Catalan", matches("catalan")) |>
	tab_spanner("English", matches("english")) |>
	tab_spanner("Spanish", matches("spanish")) |>
	tab_spanner("Degree of Exposure (%)", matches("doe")) |>
	cols_label(n = "Testing sessions",
			   n_participants = "Participants",
			   age_mean = "M (SD)",
			   age_min = "Range",
			   doe_catalan_mean = "M (SD)",
			   doe_english_mean = "M (SD)",
			   doe_spanish_mean = "M (SD)") |>
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	tab_style(cell_text(style = "italic"),
			  cells_column_spanners())
```

@@

We collected vocabulary data using parental responses to the Oxford Communicative Development Inventory (OCDI) [@hamilton2000infant]. The OCDI is an adaptation of the MacArthur-Bates Communicative Development Inventory [@fenson1994variability] to British English. The OCDI includes a vocabulary checklist containing 418 words from 21 semantic-functional categories (e.g., action words, animals, household objects, adverbs, etc.). For each word, caregivers are asked to answer if they child is able to *understand*, *understand and say* or does not understand or say the word. We calculated participants' receptive vocabulary size scores as the number of words that caregivers marked as *understands* or *understands and says*. Families were sent the questionnaire immediately after each experimental session, and were given two weeks to fill it. In the case that a complete response to the OCDI was not provided within the two-week limit, the participants' testing session was excluded from the analyses (*n* = `r vocabulary_oxf_exc`). @fig-vocabulary-oxf shows the distribution of participants' vocabulary sizes across ages.

```{r fig-vocabulary}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles. Vocabulary size scores from the recurrent participants are shown linked. "
#| fig-width: 6
#| fig-height: 4
vocab_plot_data <- vocabulary |>
	inner_join(filter(attrition_participants, is_valid_participant),
			   by = join_by(session_id)) |>
	inner_join(select(participants, location, lp, age,
					  child_id, session_id, age_group),
			   by = join_by(child_id, session_id)) |> 
	filter(is_valid_participant) |> 
	select(child_id, session_id, vocab_id, is_imputed, l1_count,
		   lp, location, age_group, age) |> 
	mutate(location = factor(location, 
							 levels = c("Oxford", "Barcelona")),
		   lp = gsub(" \\(English\\)", "", lp),
		   side = case_when(location=="Oxford" ~ "both",
		   				 lp=="Monolingual" ~ "left",
		   				 lp=="Bilingual" ~ "right"),
		   study = if_else(location=="Oxford", 
		   				"Study 1 (English)", "Study 2 (Catalan/Spanish)"))

vocab_plot_means <- vocab_plot_data |> 
	summarise(l1_count_sd = sd(l1_count, na.rm = TRUE),
			  l1_count = mean(l1_count, na.rm = TRUE),
			  n = n(),
			  .by = c(age, side, lp, study)) |> 
	mutate(.lower = l1_count - (l1_count_sd/sqrt(n)),
		   .upper = l1_count + (l1_count_sd/sqrt(n)))

vocab_plot_data |> 
	filter(study=="Study 1 (English)") |> 
	ggplot(aes(age, l1_count, 
			   side = side)) +
	geom_point(data = filter(vocab_plot_means, 
							 study=="Study 1 (English)"), 
			   colour = clrs[1],
			   show.legend = FALSE,
			   size = 2,
			   shape = 1,
			   stroke = 1) +
	geom_smooth(method = "lm", 
				formula = "y ~ x",
				colour = clrs[1],
				fill = clrs[1]) +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(limits = c(0, 575),
					   breaks = seq(0, 1e3, 50)) +
	scale_x_continuous(breaks = seq(0, 50, 2)) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  axis.title.x = element_blank(),
		  legend.title = element_blank())
```

### Design

Participants were presented with 32 trials in random order, which belonged to three conditions: *Unrelated* (16 trials), *Related/Non-cognate* (8 trials), and *Related/Cognate* (8 trials). In *Unrelated* trials, target was phonologically unrelated with the English and Spanish labels of the prime picture (e.g., $\text{door}_{\text{ENG}}/\text{puerta}_{\text{SPA}}-\text{sock}_{\text{ENG}}$). In *Related/Non-cognate* trials, the target shared phonological onset with the English prime label, but not with the Spanish prime label (e.g., $\textbf{t}\text{ree}_{\text{ENG}}/\text{árbol}_{\text{SPA}}-\textbf{t}\text{ruck}_{\text{ENG}}$). In *Related/Cognate* trials, the target shared phonological overlap with both English and Spanish prime labels ($\textbf{b}\text{oot}_{\text{ENG}}/\textbf{b}\text{ota}_{\text{SPA}}-\textbf{b}\text{ird}_{\text{ENG}}$). Especial attention was paid to avoiding semantic or taxonomic relationships between prime and target words, and between prime and distractor words. Distractors were always phonologically unrelated to the prime and target labels in the same trial.

@fig-task illustrates the sequence of a trial. Each trial started with the presentation of an attention getter for 3,000 ms. Then, the prime picture was presented in silence in the centre of the screen for 1,500 milliseconds. Fifty milliseconds after the offset of the prime image, an auditory label was played, 700 milliseconds after the onset of the word, the target and distractor pictures were presented side-by-side during 1,000 milliseconds until the end of the trial. After this, the attention getter of the next trial was immediately presented. Each experimental session lasted approximately 10 minutes.

@@

We created four stimuli lists, to which participants were randomly allocated. Recurrent participants were always allocated to the same list. Each list contained a different subset of prime-target-distractor triples. Target-distractor pairs were kept constant across lists: all infants, regardless of the list to which they were allocated, were presented with the *ball*-*trousers* target distractor pair. Each target-distractor pair was presented twice in their corresponding lists, alternating which word acted as target, and which as distractor. For instance, the *ball*-*trousers* pair was presented in all lists twice, once with *ball* as target, and another with *trousers* as target. Across lists, the same target-distractor pair appeared with a different prime, counterbalancing the condition to which it belonged. For instance, in list one the *ball*-*trousers* pair was preceded by the prime *bike* (*Related/Cognate*), by *butterfly* (*Related/Non-cognate*) in list two, and by *star* and *nose* (*Unrelated*) in lists three and four, respectively.

@tbl-stimuli shows a detailed summary of the stimuli properties, broken down by trial type and testing language. Trials included in each condition had equivalent length (number of phonemes) and lexical frequency. Lexical frequencies were extracted from the English corpora from the CHILDES database [@macwhinney2000childes; @sanchez2019childes] as counts per million words, and transformed into Zipf scores for easier cross-language comparison [@zipf1945meaning; @van2014subtlex]. In addition to sharing phonological onset, prime and target shared higher overall phonological similarity---as measured using the Levenshtein distance---between their phonological transcriptions---in the *Related/Cognate* and the *Related/Non-cognate* conditions, compared to the *Unrelated* condition. In *Related/Cognate* trials, the prime label in the language of test shared a higher overall phonological similarity with its translation, compared to *Related/Non-cognate* trials. Since the number of cognate and non-cognate primes was balanced in *Unrelated* trials, overall phonological similarity between prime labels prime labels in the *Unrelated* condition had an intermediate overall phonological similarity. Audio durations were also equivalent across conditions.

```{r tbl-stimuli}
#| label: tbl-stimuli
#| tbl-cap: "Summary of stimuli properties by trial type."
stimuli |>
	unnest_wider(c(freq, nphon)) |>
	summarise(across(c(matches("n_phon|freq_|lv_|nphon"), duration),
					 list(	mean = \(x) mean(x, na.rm = TRUE),
					 	  sd = \(x) sd(x, na.rm = TRUE))),
			  .by = c(trial_type, test_language)) |>
	select(-matches("distractor")) |>
	mutate(test_language = if_else(test_language=="English",
								   paste0("Study 1: ", test_language),
								   paste0("Study 2: ", test_language))) |>
	arrange(test_language) |> 
	gt(rowname_col = "trial_type",
	   groupname_col = "test_language") |>
	# summary_rows(groups = c("Unrelated", "Non-cognate", "Cognate"),
	# 			 columns = is.numeric,
	# 			 fns = list(Total ~ mean(.)),
	# 			 fmt = ~ fmt_number(., decimals = 2)) |>
	grand_summary_rows(	columns = is.numeric,
						fns = list(Total ~ mean(.)),
						fmt = ~ fmt_number(., decimals = 2)) |>
	tab_spanner("# Phonemes", matches("nphon")) |>
	tab_spanner("Frequency (Zipf)", matches("freq")) |>
	tab_spanner("Phon. similarity [0-1]", matches("lv")) |>
	fmt_number(is.numeric, decimals = 1) |>
	cols_merge(c(freq_prime_mean, freq_prime_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(freq_target_mean, freq_target_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(nphon_prime_mean, nphon_prime_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(nphon_target_mean, nphon_target_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(lv_pp_mean, lv_pp_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(lv_pt_mean, lv_pt_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(duration_mean, duration_sd),
			   pattern = "{1} ({2})") |>
	cols_label(trial_type = "",
			   freq_prime_mean = "Prime",
			   freq_target_mean = "Target",
			   lv_pp_mean = "Prime-Prime",
			   lv_pt_mean = "Prime-Target",
			   duration_mean = "Duration (s)",
			   nphon_prime_mean = "Prime",
			   nphon_target_mean = "Target") |>
	tab_style(cell_text(style = "italic"),
			  list(cells_column_labels(),
			  	 cells_column_spanners(),
			  	 cells_grand_summary()))
```

The auditory stimuli were natural exemplars of the selected target words, spoken by a Southern British English female speaker who was instructed to pronounce each word in a toddler-directed manner. We used the Audacity and Praat [@boersma2001speak] software packages to trim, denoised, and normalised their amplitude. The visual stimuli were realistic photographic representations of a typical exemplars of the prime, target, and distractor words. Image backgrounds were removed from the original pictures using the GNU Image Manipulation Program (GIMP), resized to a rectangle of a maximum of 400 pixels height or wide, and finally placed in the centre of a 50% grey rectangle square of 500 $\times$ 500 pixels. The final stimuli had a resolution of 72 dpi. When presented in the eye-tracker screen, the areas of interest (AOI) occupied an area of 13.23 $\times$ 13.23 cm (11.613$^{\circ}$ visual angle from participants' perspective).

Testing took place in a sound-proof room at the BabyLab of the University of Oxford. Participants sat on their caregivers' lap in a dimly lit testing booth while the experimenter conducted the experiment from outside. Caregivers were instructed to keep their eyes shut (to avoid recording their gaze, instead of the participant's), to be still, and to avoid interacting with the participant verbally or non-verbally. Participants sat at approximately 65 cm from the eye-tracker and a XX-in screen of $1929\times1080$ screen resolution. The study was run using a custom Matlab script, PresentMate, based on the PsychToolbox-3 extension [@kleiner2007s; @brainard1997psychophysics; @pelli1997videotoolbox]. Visual fixations were recorded using a Tobii TX 300 eye-tracker and a 23-in screen of 1920 $\times$ 1080 resolution. The Tobii Analytics SDK 3.0 was used to interact with the eye-tracking while the experiment was running. Sampling rate was set at 120 Hz. A 9-point calibration was performed before every experimental session, in which the picture of a colourful beach ball was presented. We set a 55% grey background for the screen during calibration and stimuli presentation. Auditory stimuli were presented through two loudspeakers located behind the screen, one to each side. The experimenter monitored the experimental from outside the room using a centrally located video camera place above the screen. After a successful calibration the experimenter triggered the onset of the first trial. Trials were presented uninterruptedly and without intervention of the experimenter until the 32 trials were presented, or the experimental session had to be stopped because of the participant's behaviour.

@@

```{r data set_oxf}
attrition_oxf <- filter(attrition, location == "Oxford")
n_trials <- nrow(attrition_oxf)
n_sessions <- n_distinct(attrition_oxf$session_id)
n_participants <- n_distinct(attrition_oxf$child_id)
n_trials_valid <- nrow(filter(
	attrition_oxf,
	is_valid_participant,
	is_valid_trial
))
n_sessions_valid <- attrition_oxf |>
	distinct(child_id, .keep_all = TRUE) |>
	pull(is_valid_participant) |>
	sum()

@@

model_summary <- get_posterior_summary(model_fits_bcn$fit_0,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval
)
